Conoco Phillips 2020

### 2020-06-04 CP_2020.Rmd
updated data analysis; removed 911; added Bear Stearns

Our donor pool will consist of Conoco Phillips shock effects observed on September 11th, 2001, March 18th, 2008, many days in September and October of 2008, and November 27, 2014. The September 11th, 2001 and the 2008 shock effects were observed during recessions that were predicated by an inversion of the yield curve, and the November 27, 2014 shock was an OPEC induced supply side shock effect. <!--Some other notable shock effects were not included in this analysis because the conditions preceding them are not simialar to those of the March 9, 2020 shock effect, among these are the October 19, 1987, August 8th, 2011, and  August 24, 2015 shock effects ([1987 Source](https://www.nber.org/chapters/c10958.pdf), [2011 Source](https://www.latimes.com/world/la-xpm-2011-aug-06-la-fi-us-debt-downgrade-20110806-story.html), [2015 Source](https://www.nytimes.com/2015/08/25/business/dealbook/daily-stock-market-activity.html)). --> 
<!-- 
October 19, 1987: https://www.nber.org/chapters/c10958.pdf
@article{shiller1988portfolio,
  title={Portfolio insurance and other investor fashions as factors in the 1987 stock market crash},
  author={Shiller, Robert J},
  journal={NBER Macroeconomics Annual},
  volume={3},
  pages={287--297},
  year={1988},
  publisher={MIT Press}
}

@@

#### Tuesday, September 11, 2001

@@

```{r}
## September 11th (no trading from T-F); 1 day nowcast
COP_close <- COP %>% dplyr::select(COP.Close, Date) %>% 
  rename(COP_Close = COP.Close)
GSPC_close <- GSPC %>% dplyr::select(GSPC.Close, Date) %>% 
  rename(GSPC_Close = GSPC.Close)
COP_close <- merge(COP_close, GSPC_close)
COP_close <- merge(COP_close, WTI_Crude)
Y <- COP_close$COP_Close[-1]
COP_close <- data.frame(COP_close[-nrow(COP_close), ], Y)
start <- which(COP_close$Date == "2001-09-10")
start_day <- as.numeric(1:nrow(COP_close) == start + 1)
COP_close <- COP_close %>% mutate(start_day = start_day)
COP_close_small <- COP_close[(start-29):(start+1), ]
COP_close_small[, 2:5] <- COP_close_small[, 2:5] * 
  inflation_adj$dollars_2020[inflation_adj$year == 2001] 
m_COP_9_11 <- lm(Y ~ COP_Close + start_day + GSPC_Close + WTI_Close, 
                 data = COP_close_small)
alpha_9_11 <- summary(m_COP_9_11)$coef[3,1:2] 
alpha_9_11
```

@@

```{r}
m_COP_9_11_no <- lm(Y ~ COP_Close + start_day, data = COP_close_small)
alpha_9_11_no <- summary(m_COP_9_11_no)$coef[3,1:2]
alpha_9_11_no
```

#### 2008 shock effects
<!-- https://www.democracynow.org/2008/9/8/headlines -->

The first 4 major shocks of 2008 were proceeded by unprecedented failures in the financial sector that rattled markets, oil prices, and Conoco Phillips stock alike. The September 12th and September 26th shock effects were proceeded by the bailouts or failures of Lehman Brothers and Washington Mutual financial institutions. The September 8th shock effect was proceeded by Fannie Mae and Freddie Mac being placed in conservatorship.

@@

```{r}
a_estimates <- rbind( c(alpha_9_11, alpha_9_11_no),
  cbind(alpha_Sept_08, alpha_Sept_08_no),
  c(alpha_11_27_14, alpha_11_27_14_no))
a_estimates[, c(2,4)] <- a_estimates[, c(2,4)]^2
colnames(a_estimates) <- c("a_1","var_a_1","a_1_no","var_a_1_no")
rownames(a_estimates) <- c("y2001","s8y2008","s12y2008","s26y2008","y2014")
```

We now compute four shock effect estimators. We compute the adjustment estimator that averages the estimated shock effects and the inverse variance weighted adjustment estimator, these two estimators are computed with and without the September 11th, 2001 shock effect. We consider analyses without the September 11th, 2001 shock effect because the stock market was closed for the week following the terrorist attack, so immediate post shock data was not actually observed.

```{r}
a_adj <- colMeans(a_estimates[, c(1,3)])
a_adj_no911 <- colMeans(a_estimates[-1, c(1,3)])
weights <- rbind(
(1 / a_estimates[, c(2,4)])[1, ] / colSums(1 / a_estimates[, c(2,4)]),
(1 / a_estimates[, c(2,4)])[2, ] / colSums(1 / a_estimates[, c(2,4)]),
(1 / a_estimates[, c(2,4)])[3, ] / colSums(1 / a_estimates[, c(2,4)]),
(1 / a_estimates[, c(2,4)])[4, ] / colSums(1 / a_estimates[, c(2,4)]),
(1 / a_estimates[, c(2,4)])[5, ] / colSums(1 / a_estimates[, c(2,4)])
)
a_adj_IVW <- colSums(weights * a_estimates[, c(1,3)])
weights_no911 <- rbind(
(1 / a_estimates[-1, c(2,4)])[1, ] / colSums(1 / a_estimates[-1, c(2,4)]),
(1 / a_estimates[-1, c(2,4)])[2, ] / colSums(1 / a_estimates[-1, c(2,4)]),
(1 / a_estimates[-1, c(2,4)])[3, ] / colSums(1 / a_estimates[-1, c(2,4)]),
(1 / a_estimates[-1, c(2,4)])[4, ] / colSums(1 / a_estimates[-1, c(2,4)])
)
a_adj_IVW_no911 <- colSums(weights_no911 * a_estimates[-1, c(1,3)])
```

@@

```{r}
shock_effects <- rbind(a_adj, a_adj_no911, 
                       a_adj_IVW, a_adj_IVW_no911)
shock_effects
```

@@

```{r}
## Additive adjustment
additive <- colMeans(a_estimates[2:4, c(1,3)]) + a_estimates[4, c(1,3)]
additive
```

@@

```{r}
weights_additive <- rbind(
(1 / a_estimates[, c(2,4)])[1, ] / colSums(1 / a_estimates[1:4, c(2,4)]),
(1 / a_estimates[, c(2,4)])[2, ] / colSums(1 / a_estimates[1:4, c(2,4)]),
(1 / a_estimates[, c(2,4)])[3, ] / colSums(1 / a_estimates[1:4, c(2,4)]),
(1 / a_estimates[, c(2,4)])[4, ] / colSums(1 / a_estimates[1:4, c(2,4)])
)
a_adj_IVW_additive <- colSums(weights_additive * a_estimates[1:4, c(1,3)])
weights_no911_additive <- rbind(
(1 / a_estimates[2:4, c(2,4)])[1, ] / colSums(1 / a_estimates[2:4, c(2,4)]),
(1 / a_estimates[2:4, c(2,4)])[2, ] / colSums(1 / a_estimates[2:4, c(2,4)]),
(1 / a_estimates[2:4, c(2,4)])[3, ] / colSums(1 / a_estimates[2:4, c(2,4)])
)
a_adj_IVW_no911_additive <- 
  colSums(weights_no911_additive * a_estimates[2:4, c(1,3)])

@@

## Additive IVW adjustment without covariates
additive_IVW_no <- as.numeric(a_adj_IVW_additive[2] + a_estimates[5, 1])
additive_IVW_no

## Additive IVW adjustment without 911
additive_IVW_no911 <- as.numeric(a_adj_IVW_no911_additive[1] + a_estimates[5, 1])
additive_IVW_no911

## Additive IVW adjustment without 911 and covariates
additive_IVW_no911_no <- as.numeric(a_adj_IVW_no911_additive[2] + a_estimates[5, 1])
additive_IVW_no911_no
```

@@

```{r}
m_COP_03_06_20 <- lm(Y ~ COP_Close + GSPC_Close + WTI_Close, 
                     data = COP_close_small[-nrow(COP_close_small), ])
preds <- predict(m_COP_03_06_20)
summary(m_COP_03_06_20)
Yhat_nothing <- -2.430990 + 1.045492 * 47.01888 + 
  0.003063 * 3023.94 + -0.210924 * 45.90
Yhat_adj <- Yhat_nothing + shock_effects[, 1]
Yhat_adj_additive <- Yhat_nothing + additive[1]
Yhat_adj_additive_IVW <- Yhat_nothing + additive_IVW
Yhat_adj_additive_IVW_no911 <- Yhat_nothing + additive_IVW_no911

@@

## additive effect does well with IVW adjustment
Yhat_adj_additive_IVW - COP_close_small$Y[nrow(COP_close_small)]
Yhat_adj_additive_IVW_no911 - COP_close_small$Y[nrow(COP_close_small)]
```

```{r}
COP_close_small$id <- 1:nrow(COP_close_small)
mat <- cbind(COP_close_small$id[nrow(COP_close_small)], 
             c(Yhat_adj, Yhat_adj_additive, Yhat_adj_additive_IVW_no911))
colnames(mat) <- c("id", "Yhat_adj")
dat <- as.data.frame(mat)
ggplot(COP_close_small, mapping = aes(x = id, y = Y)) + 
  labs(title = "Conoco Phillips forecasting", 
       x = "Day", y = "Closing price (in USD)") +
  geom_point() + 
  geom_point(data = dat, aes(x = id, y = Yhat_adj), col = 2, pch = 2:7) + 
  geom_point(data = data.frame(x = unique(dat$id), y =Yhat_nothing), 
             aes(x = x, y = y), col = "blue", cex = 1.5) + 
  geom_line(aes(x = id, y = c(m_COP_03_06_20$fitted.values, Yhat_nothing)), 
            col = "blue") + 
    annotate("text", x=1, y=seq(from = 47, to = 38,length.out=7), 
             label=c("no adjustment","mean adjustment",
                     "mean adjustment (no 9/11)","IVW adjustment",
                     "IVW adjustment (no 9/11)", "aggregate adjustment", 
                     "aggregate IVW adjustment"), 
             hjust=0) + 
    annotate("point", x = 0, y=seq(from = 47, to = 38,length.out=7), 
             pch = c(16,2:7),
             color = c("blue","red","red","red","red","red","red")) +
  theme_minimal()
```

@@

```{r}
m_COP_03_06_20 <- lm(Y ~ COP_Close, 
                     data = COP_close_small[-nrow(COP_close_small), ])
summary(m_COP_03_06_20)
preds <- predict(m_COP_03_06_20)
Yhat_nothing_no <- -1.25682 + 1.01170 * 47.00149
Yhat_adj_no <- Yhat_nothing_no + shock_effects[, 2]
Yhat_adj_additive_no <- Yhat_nothing_no + additive[2]
Yhat_adj_additive_IVW_no <- Yhat_nothing + additive_IVW_no
Yhat_adj_additive_IVW_no911_no <- Yhat_nothing + additive_IVW_no911_no

@@

## additive effect does well with IVW adjustment
Yhat_adj_additive_IVW_no - COP_close_small$Y[nrow(COP_close_small)]
Yhat_adj_additive_IVW_no911_no - COP_close_small$Y[nrow(COP_close_small)]
```

```{r}
COP_close_small$id <- 1:nrow(COP_close_small)
mat <- cbind(COP_close_small$id[nrow(COP_close_small)], 
             c(Yhat_adj,  Yhat_adj_additive_no, Yhat_adj_additive_IVW_no911_no))
colnames(mat) <- c("id", "Yhat_adj")
dat <- as.data.frame(mat)
ggplot(COP_close_small, mapping = aes(x = id, y = Y)) + 
  labs(title = "Conoco Phillips forecasting (without covariates)", 
       x = "Day", y = "Closing price (in USD)") +
  geom_point() + 
  geom_point(data = dat, aes(x = id, y = Yhat_adj), 
                             col = 2, pch = 2:7) + 
  geom_point(data = data.frame(x = unique(dat$id), y =Yhat_nothing_no), 
             aes(x = x, y = y), col = "blue", cex = 1.5) + 
  geom_line(aes(x = id, y = c(m_COP_03_06_20$fitted.values,
                              Yhat_nothing_no)), 
            col = "blue") +
      annotate("text", x=1, y=seq(from = 47, to = 38,length.out=7), 
             label=c("no adjustment","mean adjustment",
                     "mean adjustment (no 9/11)","IVW adjustment",
                     "IVW adjustment (no 9/11)", "aggregate adjustment",
                     "aggregate IVW adjustment"), 
             hjust=0) + 
    annotate("point", x = 0, y=seq(from = 47, to = 38,length.out=7), 
             pch = c(16,2:7),
             color = c("blue","red","red","red","red","red","red")) +
  theme_minimal()
```

>>>>

Our donor pool will consist of Conoco Phillips shock effects observed on March 17th, 2008, many days in September and October of 2008, and November 27, 2014. The 2008 shock effects were observed during recessions that were predicated by an inversion of the yield curve, and the November 27, 2014 shock was an OPEC induced supply side shock effect. <!--Some other notable shock effects were not included in this analysis because the conditions preceding them are not simialar to those of the March 9, 2020 shock effect, among these are the October 19, 1987, August 8th, 2011, and  August 24, 2015 shock effects ([1987 Source](https://www.nber.org/chapters/c10958.pdf), [2011 Source](https://www.latimes.com/world/la-xpm-2011-aug-06-la-fi-us-debt-downgrade-20110806-story.html), [2015 Source](https://www.nytimes.com/2015/08/25/business/dealbook/daily-stock-market-activity.html)). --> 
<!-- 
October 19, 1987: https://www.nber.org/chapters/c10958.pdf
@article{shiller1988portfolio,
  title={Portfolio insurance and other investor fashions as factors in the 1987 stock market crash},
  author={Shiller, Robert J},
  journal={NBER Macroeconomics Annual},
  volume={3},
  pages={287--297},
  year={1988},
  publisher={MIT Press}
}

@@

#### Monday, March 17th, 2008

@@

```{r}
## March 17th; 1 day nowcast
COP_close <- COP %>% dplyr::select(COP.Close, Date) %>% 
  rename(COP_Close = COP.Close)
GSPC_close <- GSPC %>% dplyr::select(GSPC.Close, Date) %>% 
  rename(GSPC_Close = GSPC.Close)
COP_close <- merge(COP_close, GSPC_close)
COP_close <- merge(COP_close, WTI_Crude)
Y <- COP_close$COP_Close[-1]
COP_close <- data.frame(COP_close[-nrow(COP_close), ], Y)
start <- which(COP_close$Date == "2008-03-14")
start_day <- as.numeric(1:nrow(COP_close) == start)
COP_close <- COP_close %>% mutate(start_day = start_day)
COP_close_small <- COP_close[(start-30):start, ]
COP_close_small[, 2:5] <- COP_close_small[, 2:5] * 
  inflation_adj$dollars_2020[inflation_adj$year == 2008] 
m_COP_3_17 <- lm(Y ~ COP_Close + start_day + GSPC_Close + WTI_Close, 
                 data = COP_close_small)
alpha_3_17 <- summary(m_COP_3_17)$coef[3,1:2] 
alpha_3_17
```

@@

```{r}
m_COP_3_17_no <- lm(Y ~ COP_Close + start_day, data = COP_close_small)
alpha_3_17_no <- summary(m_COP_3_17_no)$coef[3,1:2] 
alpha_3_17_no
```

#### 2008 September shock effects
<!-- https://www.democracynow.org/2008/9/8/headlines -->

The first 3 major shocks in September, 2008 were proceeded by failures in the financial sector that rattled markets, oil prices, and Conoco Phillips stock alike. The September 12th and September 26th shock effects were proceeded by the bailouts or failures of Lehman Brothers and Washington Mutual financial institutions. The September 8th shock effect was proceeded by Fannie Mae and Freddie Mac being placed in conservatorship.

@@

```{r}
a_estimates <- rbind( c(alpha_3_17, alpha_3_17_no),
  cbind(alpha_Sept_08, alpha_Sept_08_no),
  c(alpha_11_27_14, alpha_11_27_14_no))
a_estimates[, c(2,4)] <- a_estimates[, c(2,4)]^2
colnames(a_estimates) <- c("a_1","var_a_1","a_1_no","var_a_1_no")
rownames(a_estimates) <- c("m2008","s8y2008","s12y2008","s26y2008","y2014")
```

We now compute shock effect estimators. We compute the adjustment estimator that averages the estimated shock effects and the inverse variance weighted adjustment estimator.

```{r}
a_adj <- colMeans(a_estimates[, c(1,3)])
weights <- rbind(
(1 / a_estimates[, c(2,4)])[1, ] / colSums(1 / a_estimates[, c(2,4)]),
(1 / a_estimates[, c(2,4)])[2, ] / colSums(1 / a_estimates[, c(2,4)]),
(1 / a_estimates[, c(2,4)])[3, ] / colSums(1 / a_estimates[, c(2,4)]),
(1 / a_estimates[, c(2,4)])[4, ] / colSums(1 / a_estimates[, c(2,4)]),
(1 / a_estimates[, c(2,4)])[5, ] / colSums(1 / a_estimates[, c(2,4)])
)
a_adj_IVW <- colSums(weights * a_estimates[, c(1,3)])
```

@@

```{r}
shock_effects <- rbind(a_adj, a_adj_IVW)
shock_effects
```

@@

```{r}
## Additive adjustment
additive <- colMeans(a_estimates[1:4, c(1,3)]) + a_estimates[5, c(1,3)]
additive
```

@@

```{r}
weights_additive <- rbind(
(1 / a_estimates[, c(2,4)])[1, ] / colSums(1 / a_estimates[1:4, c(2,4)]),
(1 / a_estimates[, c(2,4)])[2, ] / colSums(1 / a_estimates[1:4, c(2,4)]),
(1 / a_estimates[, c(2,4)])[3, ] / colSums(1 / a_estimates[1:4, c(2,4)]),
(1 / a_estimates[, c(2,4)])[4, ] / colSums(1 / a_estimates[1:4, c(2,4)])
)
a_adj_IVW_additive <- colSums(weights_additive * a_estimates[1:4, c(1,3)])

@@

## Additive IVW adjustment without covariates
additive_IVW_no <- as.numeric(a_adj_IVW_additive[2] + a_estimates[5, 1])
additive_IVW_no
```

@@

```{r}
m_COP_03_06_20 <- lm(Y ~ COP_Close + GSPC_Close + WTI_Close, 
                     data = COP_close_small[-nrow(COP_close_small), ])
preds <- predict(m_COP_03_06_20)
summary(m_COP_03_06_20)
Yhat_nothing <- -2.430990 + 1.045492 * 47.01888 + 
  0.003063 * 3023.94 + -0.210924 * 45.90
Yhat_adj <- Yhat_nothing + shock_effects[, 1]
Yhat_adj_additive <- Yhat_nothing + additive[1]
Yhat_adj_additive_IVW <- Yhat_nothing + additive_IVW

@@

## additive effect does well with IVW adjustment
Yhat_adj_additive_IVW - COP_close_small$Y[nrow(COP_close_small)]
```

Here is a plot without additive adjustment estimators included:

```{r}
COP_close_small$id <- 1:nrow(COP_close_small)
mat <- cbind(COP_close_small$id[nrow(COP_close_small)], 
             c(Yhat_adj))
colnames(mat) <- c("id", "Yhat_adj")
dat <- as.data.frame(mat)
ggplot(COP_close_small, mapping = aes(x = id, y = Y)) + 
  labs(title = "Conoco Phillips forecasting", 
       x = "Day", y = "Closing price (in USD)") +
  geom_point() + 
  geom_point(data = dat, aes(x = id, y = Yhat_adj), col = 2, pch = 2:3) + 
  geom_point(data = data.frame(x = unique(dat$id), y =Yhat_nothing), 
             aes(x = x, y = y), col = "blue", cex = 1.5) + 
  geom_line(aes(x = id, y = c(m_COP_03_06_20$fitted.values, Yhat_nothing)), 
            col = "blue") + 
    annotate("text", x=1, y=seq(from = 45, to = 40, length.out=3), 
             label=c("no adjustment","mean adjustment","IVW adjustment"), 
             hjust=0) + 
    annotate("point", x = 0, y=seq(from = 45, to = 40, length.out=3), 
             pch = c(16, 2:3),
             color = c("blue","red","red")) +
  theme_minimal()
```

Here is a plot with additive adjustment estimators included:

```{r}
COP_close_small$id <- 1:nrow(COP_close_small)
mat <- cbind(COP_close_small$id[nrow(COP_close_small)], 
             c(Yhat_adj, Yhat_adj_additive, Yhat_adj_additive_IVW))
colnames(mat) <- c("id", "Yhat_adj")
dat <- as.data.frame(mat)
ggplot(COP_close_small, mapping = aes(x = id, y = Y)) + 
  labs(title = "Conoco Phillips forecasting", 
       x = "Day", y = "Closing price (in USD)") +
  geom_point() + 
  geom_point(data = dat, aes(x = id, y = Yhat_adj), col = 2, pch = 2:5) + 
  geom_point(data = data.frame(x = unique(dat$id), y =Yhat_nothing), 
             aes(x = x, y = y), col = "blue", cex = 1.5) + 
  geom_line(aes(x = id, y = c(m_COP_03_06_20$fitted.values, Yhat_nothing)), 
            col = "blue") + 
    annotate("text", x=1, y=seq(from = 47, to = 40,length.out=5), 
             label=c("no adjustment","mean adjustment",
                     "IVW adjustment", "aggregate adjustment", 
                     "aggregate IVW adjustment"), 
             hjust=0) + 
    annotate("point", x = 0, y=seq(from = 47, to = 40,length.out=5), 
             pch = c(16,2:5),
             color = c("blue","red","red","red","red")) +
  theme_minimal()
```

@@

```{r}
m_COP_03_06_20 <- lm(Y ~ COP_Close, 
                     data = COP_close_small[-nrow(COP_close_small), ])
summary(m_COP_03_06_20)
preds <- predict(m_COP_03_06_20)
Yhat_nothing_no <- -1.25682 + 1.01170 * 47.00149
Yhat_adj_no <- Yhat_nothing_no + shock_effects[, 2]
Yhat_adj_additive_no <- Yhat_nothing_no + additive[2]
Yhat_adj_additive_IVW_no <- Yhat_nothing + additive_IVW_no

@@

## additive effect does well with IVW adjustment
Yhat_adj_additive_IVW_no - COP_close_small$Y[nrow(COP_close_small)]
```

```{r}
COP_close_small$id <- 1:nrow(COP_close_small)
mat <- cbind(COP_close_small$id[nrow(COP_close_small)], 
             c(Yhat_adj_no,  Yhat_adj_additive_no, Yhat_adj_additive_IVW_no))
colnames(mat) <- c("id", "Yhat_adj")
dat <- as.data.frame(mat)
ggplot(COP_close_small, mapping = aes(x = id, y = Y)) + 
  labs(title = "Conoco Phillips forecasting (without covariates)", 
       x = "Day", y = "Closing price (in USD)") +
  geom_point() + 
  geom_point(data = dat, aes(x = id, y = Yhat_adj), 
                             col = 2, pch = 2:5) + 
  geom_point(data = data.frame(x = unique(dat$id), y =Yhat_nothing_no), 
             aes(x = x, y = y), col = "blue", cex = 1.5) + 
  geom_line(aes(x = id, y = c(m_COP_03_06_20$fitted.values,
                              Yhat_nothing_no)), 
            col = "blue") +
      annotate("text", x=1, y=seq(from = 47, to = 40,length.out=5), 
             label=c("no adjustment","mean adjustment","IVW adjustment",
                     "aggregate adjustment","aggregate IVW adjustment"), 
             hjust=0) + 
    annotate("point", x = 0, y=seq(from = 47, to = 40,length.out=5), 
             pch = c(16,2:5),
             color = c("blue","red","red","red","red")) +
  theme_minimal()
```
